Gray Box Optimization using the Parameter-less Population Pyramid

### 2015-01-12 Gray Box Optimization using the Parameter-less Population Pyramid/main.tex
Gray Box up to results

\section{Hamming Ball Hill Climbing}
The first algorithm to rigorously exploit the features of the gray box domain
described in Section~\ref{sec-gray-box} was \cite{whitley:2013:greedy}. In this study
a hill climber was developed which found the approximate next best improving move in \BigO{1} time.
This is achieved by first determining the fitness effect $delta$ of making each move, and then updating
only effected move each time a change to the solution is made. Due to the requirements of the domain,
at most \BigO{ck} moves can be effected, with each requiring at most \BigO{1} subfunction calls
to update their $delta$. Combined with move binning, this technique is able to perform both
first improvement and next best hill climbing in \BigO{I+N}, where $I$ is the number of improving
moves required to reach a local optimum. In comparison, the black box approach for these searches can require
\BigO{IN^2}.

@@

TODO Overall Time Complexity

\section{Experimentation}
\label{sec-experiments}
~\cite{wright:2000:solvingnk}

\subsection{The Effect of Radius}
TODO Plot for single K value with x=radius, y=fitness, color=solver

>>>>

\section{Hamming Ball Hill Climbing}
\label{sec-hamming}
The first algorithm to rigorously exploit the features of the gray box domain
described in Section~\ref{sec-gray-box} was \cite{whitley:2013:greedy}. In this study
a hill climber was developed which found the approximate next best improving move in \BigO{1} time.
This is achieved by first determining the fitness effect $delta$ of making each move, and then updating
only effected move each time a change to the solution is made. Due to the requirements of the domain,
at most \BigO{ck} moves can be effected, with each requiring at most \BigO{1} subfunction calls
to update their $delta$. Combined with move binning, this technique is able to perform both
first improvement and next best hill climbing in \BigO{I+N}, where $I$ is the number of improving
moves required to reach a local optimum. In comparison, the black box approach for these searches can require
\BigO{IN^2}.

@@

\section{Experimentation}
\label{sec-experiments}
To compare these gray box optimization techniques we have chosen
NKq-Landscapes~\cite{chicano:2014:ball}. NKq-Landscapes create
a collection of randomly generated problem instances given a
higher level problem class description. Each instance is described
by a series of $N$ subfunctions, each corresponding to a variable
in the solution. This subfunction uses its variable and $K$ other
variables in the solution to calculate a fitness value. Fitness values
are represented as a randomly generated lookup table, such that table
entries are integers in the range $[0..q-1]$. As each subfunction reads
$K+1$ variables, the table's size and $q$ are set to $2^{K+1}$. The quality
of a solution is equal to the sum of the values returned by these subfunctions.

In this work we shall consider two methods for choosing the $K$ variables
each $x_i$'s subfunction depends on: Nearest Neighbor NKq and Unrestricted NKq.
In Nearest Neighbor NKq each variable depends on the $K$ which immediately follow
it in the genome, with variables at the end of the solution wrapping to depend on the
starting variables. Landscapes of this form can be solved in polynomial
time~\cite{wright:2000:solvingnk}, allowing comparisons of how quickly each optimization
algorithm can find the global optima. Nearest Neighbor NKq also ensures that $c=k=K+1$
and that both $c$ and $k$ do not increase as $N$ increases, ensuring that the efficiency
conclusions made in Section~\ref{sec-hamming} are applicable.

Unrestricted NKq landscapes draw the $K$ dependencies at random. For $K > 1$ it is
NP-Hard to find the global optimum of these landscapes. This also means that while
$k$ remains fixed, the maximum number of subfunctions a variable appears in ($c$) can increase
as $N$ increases. As a result some of the efficiency claims in Section~\ref{sec-hamming}
may not be applicable.

For each $N$ and $K$ tested, we generated 50 Nearest Neighbor NKq and Unrestricted NKq instances.
Each method was run once on each instance, and limited to 3 hours of computation and 4 GB of memory.
Runs were performed on 2.5GHz Intel Xeon E5-2670v2 processors using the C++11 code provided
from our website.\footnote{\url{https://github.com/name_of_repository/anonymous_name_of_project}}
Each time the run achieved a new best fitness we record current amount of processing time used.
Timing includes the discovery of subgraphs to allow for comparison between different radius values.
When reporting the ``best'' fitness for an instance we mean the best fitness found by any method before
the time limit is reached. On all Nearest Neighbor NKq instances the ``best'' fitness
is the global optimum, verified using dynamic programming.

\subsection{The Effect of Radius}
The only algorithm parameter in the Hamming Ball Hill Climber, TUX, and Gray Box P3
is the radius of the hamming ball.

TODO Plot for single K value with x=radius, y=fitness, color=solver
